---
title: "Data analysis pipeline for the paper 'The Link Between Flow and Performance is Moderated by Task Experience'"
author: "Jussi Palomäki, Tuisku Tammi, Benjamin Ultan Cowley"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 2
    theme: united
---

```{r combine_data pipeline chunk, include=FALSE}
# attach packages
library(magrittr)
library(tidyverse)
library(data.table)
library(stringr)
library(purrr)
library(broom)
library(jsonlite)
library(DT)
library(here)


source(file.path(here(), 'R', 'utils.R'))
# set folder with data files
# 
indir <- file.path(here(), '..', 'data')
# Also do something with blobs and steps (large tables)
BIGDATA <- FALSE
if (BIGDATA){
  t <- c("blob", "run", "step")
}else{
  t <- "run"
}


### BACKRGOUND DATA ----
# bkgd <-
#   read.csv(file.path(indir, "background_2019.csv")) %>%
#   mutate(ID = ID + 9) %>%
#   rbind(read.csv(file.path(indir, "background_2017.csv")), .)
# head(bkgd)
background19 <- read.csv(file.path(indir, "background_2019.csv"))


### Behavioral (CogCarSim) data ----
# import the cogcarsim2 databases
tb17 <- readCCSdb(file.path(indir, "cogcarsim2_2017.db"), t, 17)
tb19 <- readCCSdb(file.path(indir, "cogcarsim2_2019.db"), t, 19)

tbl_runs <- rbind(tb17$tbl_run, tb19$tbl_run)
head(tbl_runs)

# remove extra (test) runs
tbl_runs %<>% # %<>% operator overwrites dataframe
  filter(startsWith(participant, '0'))

if (BIGDATA){
  tbl_blob <- rbind(tb17$tbl_blob, tb19$tbl_blob)
  head(tbl_blob)
  tbl_step <- rbind(tb17$tbl_step, tb19$tbl_step)
  head(tbl_step)

  # filter other tables with updated runs
  tbl_blob %<>%
    filter(run %in% tbl_runs$run)
  
  tbl_step %<>%
    filter(run %in% tbl_runs$run)
  
}
rm(tb17, tb19)


# clean up / modify behavioural data
game_data <- tbl_runs %>%
  dplyr::select(-run) %>%
  # extract old participant variable into new participant, session, run variables
  tidyr::extract(participant, c('participant', 'session', 'run'),
                 "([[:alnum:]]+)-([[:alnum:]]+)-([[:alnum:]]+)") %>%
  mutate_at(vars(participant,session,run), as.numeric) %>%
  # this is a pure hack to change 2019 participant numbers to be 10:18
  mutate(participant = participant + ((provenance - 17) * 4.5)) %>%
  # arrange rows to create cumulative run variable
  arrange(participant, session, run) %>%
  # create ID = integer factor for participants
  mutate(ID = as.factor(as.integer(participant))) %>%
  # mutate participant variable into the right format (1 -> "01")
  mutate(participant = formatC(participant, width = 2, format = "d", flag = "0")) %>%
  group_by(participant) %>%
  # also log(duration) and log(cumrun) variables for plotting
  dplyr::mutate(cumrun = row_number(),
         ln.duration = log(duration),
         ln.cumrun = log(cumrun)) %>% 
  ungroup # remember to ungroup!

head(game_data)



### read Flow data ----
fss <-
  fread(file.path(indir, "fss_data_2019.csv")) %>%
  mutate(participant = participant + 9) %>%
  rbind(fread(file.path(indir, "fss_data_2017.csv")), .)
head(fss) # Flow + importance items Q1-Q13, skill-demand items A1-A3 (not used here)

# indicate fss questions for flow factors: fluency, absorption, flow, perceived importance
fluency_items <- c('Q2', 'Q7', 'Q8', 'Q9', 'Q4', 'Q5')
abs_items <- c('Q6', 'Q10') # NOTE THAT 1) Q1 has been removed, 2) items Q4 and Q5 now load onto absorption, due to theoretical and cfa issues!
flow_items <- c(fluency_items, abs_items)
pi_items <- c('Q11', 'Q12', 'Q13')
additional_items <- c('A1', 'A2', 'A3') #additional items! (line added 8.7.20)
#pi_items <- c('Q11', 'Q12')


# compute flow factors per run
fss_items <- fss %>%
  mutate(fluency = rowMeans(dplyr::select(., all_of(fluency_items))),
         absorption = rowMeans(dplyr::select(., all_of(abs_items))),
         flow = rowMeans(dplyr::select(., all_of(flow_items))),
         pi_total = rowMeans(dplyr::select(., all_of(pi_items)))) %>%
  # add also total perceived importance (mean)
  dplyr::rename(pi1 = Q11,
         pi2 = Q12,
         pi3 = Q13) %>%
  dplyr::rename(comp1 = A1,
         comp2 = A2,
         comp3 = A3) %>% ##additional items on perceived competence! (line added 8.7.20)
  dplyr::select(participant:run, fluency, absorption, flow, pi1:pi3, comp1:comp3, pi_total) %>%
  #dplyr::select(participant:run, fluency, absorption, flow, pi1:pi3, pi_total) %>%
  # create ID = integer factor for participants
  mutate(ID = as.factor(as.integer(participant))) %>%
  # mutate participant variable into the right format (1 -> "01")
  mutate(participant = formatC(participant, width = 2, format = "d", flag = "0"))

head(fss_items)


# join game data and fss data into a single dataframe
fss_game <- game_data %>%
  dplyr::select(participant:run, ID, collisions, duration, ln.duration, distance, cumrun, ln.cumrun) %>%
  left_join(fss_items, by = c('participant', 'session', 'run', 'ID')) # if no vars given, use the ones with identical names

# see if everything is as it should be (e.g. no leftover rows of data, sane amount of sessions etc.)
summary(fss_game)



### Extracting learning curve coefficients and predicted learning curves ----
# Fit a log-log regression curve for each participant separately: 
#     log(duration) ~ log(cumulative run)
# to get slope and intercept coefficients
fss_learning <- fss_game %>%
  dplyr::group_by(participant) %>%
  nest() %>% 
  # fit a linear model using log-log formula on durations and cumruns per participant
  dplyr::mutate(fit = purrr::map(data, ~lm(ln.duration ~ ln.cumrun, data = .)), # make models and... (note: maps-library masks map() so I added purrr::)
       coef = purrr::map(fit, tidy), # ...get coefficients and...
       data_aug = purrr::map(fit, augment)) %>% #...get predictions
  unnest(coef) %>%
  # predicted curve slope and interception
  dplyr::select(participant, term, estimate, data_aug) %>%
  spread(term, estimate) %>%
  dplyr::rename(slope = ln.cumrun,
       intercept = `(Intercept)`) %>%
  unnest(data_aug) %>%
  ungroup %>%
  # join new ones to fss_game -> fss_learning
  dplyr::select(participant:.fitted, intercept, slope) %>%
  dplyr::rename(learning_curve = .fitted) %>%
  left_join(fss_game, by = c('participant', 'ln.duration', 'ln.cumrun')) %>%
  dplyr::select(participant, session, run, everything()) %>%
  # acquire  deviation from predicted curve
  dplyr::mutate(deviation = ln.duration - learning_curve) %>%
  group_by(participant) %>%
  # acquire flow z-scores
  dplyr::mutate(z.flow = ((flow-mean(flow)) / sd(flow))) %>%
  ungroup()

# spit out newly created data
head(fss_learning)

###
###
###

outgoners <- function(x, var, thr=2.5){
  # remove z-score outliers of a given variable at some multiple of sigma
  tz <- scale(var)
  outgoners <- x[tz<thr,]
}


chkldpkg <- function(x){
  for( i in x ){
    #  require returns TRUE invisibly if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , dependencies = TRUE )
      #  Load package after installing
      require( i , character.only = TRUE )
    }
  }
}

detach_package <- function(pkg, character.only = FALSE)
{
  if(!character.only)
  {
    pkg <- deparse(substitute(pkg))
  }
  search_item <- paste("package", pkg, sep = ":")
  while(search_item %in% search())
  {
    detach(search_item, unload = TRUE, character.only = TRUE)
  }
}

zdiff <- function(lm1, lm2){
  # test the slopes of two linear models for significant difference
  # the model IV should be after Intercept - 2nd in the coefficients list
  slmF <- summary(lm1)
  slmL <- summary(lm2)
  zdiff <- (slmF$coefficients[2,1] - slmL$coefficients[2,1]) / 
    sqrt((slmF$coefficients[2,1]*slmF$coefficients[2,2])^2 + 
           (slmL$coefficients[2,1]*slmL$coefficients[2,2])^2)
}


#bootstrapped CIs for differences
bootstrapCIs <- function( df, varname, runs ){
  require( boot )
  set.seed(42)
  boot.fun <- function( x, ind ){
    c("mean" = mean(x[ind]), "median" = median(x[ind]), "var" = var(x[ind]))    
  }
  
  if (nrow(df) > 1){
    b <- boot( df[[varname]], boot.fun, runs)
    bci <- boot.ci(b, type = "bca")
  }else{
    bci <- list(bca = list(conf = 1, 1, 0, df[[varname]], df[[varname]]))
  }
}

bootin <- function(df, varname){
  bCI <- bootstrapCIs(df, varname, 5000)
  data.frame(bCI$bca)
}

quadfit <- function(df, dependent, factor){

  df$quads <- fitted(lm(df[[dependent]]~factor+I(factor^2)))
  df
}

#define function that removes NA before calculating the cosine similarity
library(lsa)
cos.na <- function(x,y)
{
  cosine(na.omit(cbind(x,y)))
}

#...same for angle
angle.na <- function(x,y){
  tmp <- na.omit(cbind(x,y))
  x <- tmp[,1]
  y <- tmp[,2]
  dot.prod <- x%*%y 
  norm.x <- norm(x,type="2")
  norm.y <- norm(y,type="2")
  theta <- acos(dot.prod / (norm.x * norm.y))
  as.numeric(theta)
}


#function to average Pearson correlations, from Alexander (1990), Bull Psych Soc, 28 (4). Formula 5
#fails for small samples n<4, and works badly for rank correlations, due to large likelihood of r = +-1
agg.pearson.cors <- function(rs.ns)
{
  na.idx <- is.na(rs.ns[,1])
  rs <- rs.ns[!na.idx,1]
  ns <- rs.ns[!na.idx,2]
  k <- length(rs)
  ns.k <- sum(ns) - k
  r.curly <- rs + (rs * (1 - rs ^ 2)) / (2 * (ns - 3))
  
  aggcors <- sum((ns - 1) * r.curly) / ns.k
  return(aggcors)
}

#function computes aggregated rank correlation of multiple samples,
agg.rank.cors <- function(rs.ns)
{
  na.idx <- is.na(rs.ns[,1]) | is.na(rs.ns[,2]) # NA index
  rs <- rs.ns[!na.idx,1]
  ns <- rs.ns[!na.idx,2]
  
  aggcors <- agg.cors(rs, ns)
  
  return(aggcors)
}


# function to aggregate correlations by Hunter-Schmidt method,
# see Zhang & Wang (2014), Multivariate Behavioral Research, 49:2, 130-148
agg.cors <- function(rs, ns)
{
  rhat <- sum(rs * ns)
  
  return( rhat / sum(ns) )
}

# GM means due to here: https://stackoverflow.com/questions/2602583/geometric-mean-is-there-a-built-in
# geometric mean, handling NA and zeros
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

# geometric mean, handling NA and optionally propagating zeros
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

center_scale <- function(df, na.rm = TRUE){
  
  if (length(df[!is.na(df)]) == 1){
    df[!is.na(df)] = 0
    return(df)
  }
  mx_df <- max(df, na.rm = na.rm)
  mn_df <- min(df, na.rm = na.rm)
  half_dist <- (mx_df - mn_df) / 2
  cs_df <- (df - (mn_df + half_dist)) / half_dist
  
  return(cs_df)
}

prettyPrintModelFit <- function(text, model){
  require(broom)
  tmp <- capture.output(glance(model))
  cat(text, "\n", tmp[1], "\n", tmp[2])
}



## ----
# Functions to generate Maximum Width Envelope analysis
# from K. Puolamäki
#
#

#' Priority queue
#' @param p order of column, i.e., output of order (x)
#' @return Lexical closure which contains the following O(1) methods:
#'  $first(), $second(), $last(), $last2(), and $remove(i) as documented
#'  in Korpela et al. (2014), appendix A.
#' @export
PQ <- function(p) {
  n <- length(p)
  idx <- 1+order(p)  # index at which item i can be found
  p <- c(NA,p,NA)      # head is at p[1] and tail at p[n+2]
  nxt <- c(2:(n+2),NA) # pointer to next
  prv <- c(NA,1:(n+1)) # pointer to previous
  first <- function() p[nxt[1]]
  second <- function() p[nxt[nxt[1]]]
  last <- function() p[prv[n+2]]
  last2 <- function() p[prv[prv[n+2]]]
  history <- rep(NA,n)
  pos <- maxpos <- 0
  remove <- function(i) {
    pos <<- maxpos <<- pos+1 # update position
    history[pos] <<- i       # add this to history
    j <- idx[i]
    prv[nxt[j]] <<- prv[j] # previous of the next is previous of the current 
    nxt[prv[j]] <<- nxt[j] # next of the previous is next of the current
    pos
  }
  ## remove/unremove previously removed items. This is not really needed
  ## but costs nothing to include here at this stage...
  goto <- function(newpos) { 
    if(newpos<pos) { # go backward
      for(i in pos:(newpos+1)) {
        j <- idx[history[i]]
        prv[nxt[j]] <<- j # unremove
        nxt[prv[j]] <<- j # unremove
      }
    } else if(newpos>pos) { # go forward
      for(i in pos:(newpos-1)) {
        j <- idx[history[i]]
        prv[nxt[j]] <<- prv[j] # remove
        nxt[prv[j]] <<- nxt[j] # remove
      }
    }
    pos <<- newpos
    pos
  }
  show <- function() list(idx=idx,p=p,nxt=nxt,prv=prv,pos=pos,
                          history=if(maxpos==0) c() else history[1:maxpos]) # for debugging
  list(first=first,second=second,last=last,last2=last2,remove=remove,goto=goto,show=show)
}

#' Finds MWE efficiently using separate validation data as documented in Korpela et al.
#' (2014). The algorithm has time complexity of O(m*n*log(n)), where m is the length of 
#' time series and n is the rows in the training data or validation data, whichever is
#' larger.
#' @param max_k integer, maximum value of k searched
#' @param max_alpha real number, fraction of validation data samples out of MWE
#' @param data_tr nXm matrix, training data
#' @param data_va n'Xm matrix, validation data
#' @return Returns a list that contains number of time series removed k, final alpha 
#'  (should be the largest alpha which is at most max_alpha), and lower and upper 
#'  bounds.
#' @export
find_mwe <- function(max_k,max_alpha,data_tr,data_va) {
  m <- dim(data_tr)[2]
  # use priority queues for both training and validation set. This finds it efficient
  # to find the largest and smallest non-removed curves.
  q_tr <- apply(data_tr,2,function(x) PQ(order(x))) 
  q_va <- apply(data_va,2,function(x) PQ(order(x)))
  lo0 <- up0 <- NULL
  k <- alpha <- -1
  removed <- 0
  while(k<max_k && removed/dim(data_va)[1]<=max_alpha) {
    idx <- matrix(c(rep(1:m,2),
                    sapply(q_tr,function(q) q$first()), sapply(q_tr,function(q) q$last()),
                    sapply(q_tr,function(q) q$second()),sapply(q_tr,function(q) q$last2())),2*m,3)
    lo <- data_tr[idx[  1:m ,c(2,1)]] # current lower envelope
    up <- data_tr[idx[-(1:m),c(2,1)]] # current upper envelope
    for(i in 1:m) { # column/time index i
      j <- q_va[[i]]$first() # index of the smallest item in validation set
      while(data_va[j,i]<lo[i] && removed/dim(data_va)[1]<=max_alpha) { 
        ## remove if below lower limit
        sapply(q_va,function(q) q$remove(j))
        j <- q_va[[i]]$first()
        removed <- removed+1
      }
      j <- q_va[[i]]$last() # index of the largest item in validation set
      while(up[i]<data_va[j,i] && removed/dim(data_va)[1]<=max_alpha) { 
        ## remove if above upper limit
        sapply(q_va,function(q) q$remove(j))
        j <- q_va[[i]]$last()
        removed <- removed+1
      }
    }
    if(removed/dim(data_va)[1]<=max_alpha) {
      lo0 <- lo
      up0 <- up
      k <- k+1
      alpha <- removed/dim(data_va)[1]
      
      ## greedy MWE algorithm: remove time series which decreases the MWE most:
      a <- aggregate(x=data.frame(gain=abs(data_tr[idx[,c(2,1)]]-data_tr[idx[,c(3,1)]])),
                     by=list(i=idx[,2]),
                     FUN=sum)
      j <- a[which.max(a[,"gain"]),"i"]
      sapply(q_tr,function(q) q$remove(j))
    }
  }
  list(k=k,alpha=alpha,lo=lo0,up=up0)
}

readdataset <- function(filename) {
  data <- read.csv(filename,header=FALSE,sep=";",na.strings="NA")[,-1]
  if(any(apply(data,2,function(x) any(is.na(x)) & any(!is.na(x))))) stop("readdataset: NAs detected")
  if(dim(data)[1]!=41) stop("readdataset: wrong dim")
  idx <- which(apply(data,2,function(x) all(!is.na(x))))
  list(data=t(as.matrix(data[,idx])),idx=idx)
}

findcurves <- function(data,mtr=5000,mva=5000,alpha=0.05,signflip=FALSE) {
  n <- dim(data)[1]
  m <- dim(data)[2]
  if(signflip) {
    samples_tr <- t(replicate(mtr,colMeans((sample(c(-1,1),size=n,replace=TRUE) %o% rep(1,m))*data)))
    samples_va <- t(replicate(mva,colMeans((sample(c(-1,1),size=n,replace=TRUE) %o% rep(1,m))*data)))
  } else {
    samples_tr <- t(replicate(mtr,colMeans(data[sample.int(n,replace=TRUE),]))) # training set
    samples_va <- t(replicate(mva,colMeans(data[sample.int(n,replace=TRUE),]))) # validation set
  } 
  a <- find_mwe(mtr-2,alpha,samples_tr,samples_va)
  q <- apply(samples_tr,2,function(x) quantile(x,probs=c(alpha/2,1-alpha/2)))
  data.frame(mean0=colMeans(data),lo=a$lo,up=a$up,k=a$k,alpha=a$alpha,lo0=q[1,],up0=q[2,])
}


```

```{r setup, include=FALSE}
# attach packages
library(easypackages)
libraries("ggplot2", "gridBase", "colorRamps", "grDevices", "lme4", "effects",
          "multcomp", "tidyverse", "lmerTest", "ggpubr", "MuMIn", "lattice",
          "robustlmm", "grid", "sjPlot", "dynlm", "viridis", "psych", "lavaan",
          "GPArotation", "lavaanPlot", "jtools", "interactions", "XML",
          "semPlot", "GGally", "ggbiplot", "gghalves", "bmlm", "mlm") # see https://link.springer.com/article/10.3758/s13428-017-0980-9#Sec10

# set working directory, i.e. folder with data files
knitr::opts_knit$set(root.dir = normalizePath(file.path(here::here(), '..', 'data')))

# create output folder?
odir <- file.path(here::here(), 'figures')
dir.create(odir, showWarnings = FALSE)
```

# Hypothesis 1a and 1b
```{r H1 - replication, message=FALSE, warning=FALSE}
```

## Hypothesis 1a and 1b analysis
```{r H1 - analysis, message=FALSE, warning=FALSE}
############
#HYPOTHESIS 1A, ANALYSIS: Replication of results of Cowley et al. (2019), effect of learning, power-law curve
############

H1a_lmer <- lmer(ln.duration ~ ln.cumrun + (ln.cumrun|participant), data=subset(fss_learning, as.numeric(ID) > 9)) #NOTE that this way of subsetting no longer works correctly when the ID factor levels are reordered (as will be done later in the pipeline!)
summary(H1a_lmer) # model summary
r.squaredGLMM(H1a_lmer) #model marginal and conditional pseudo-r2
plot(H1a_lmer) # model diagnosticsi
qqnorm(residuals(H1a_lmer)) #qq-plot
qqline(residuals(H1a_lmer)) #line of "perfect normality"

############
#HYPOTHESIS 1B, ANALYSIS: The f~d effect
############

H1b_lmer <- lmer(flow ~ deviation + (deviation|participant), 
                        data=filter(fss_learning, as.numeric(ID) > 9))
summary(H1b_lmer) # model summary
r.squaredGLMM(H1b_lmer) #model marginal and conditional pseudo-r2
plot(H1b_lmer) # model diagnostics
qqnorm(residuals(H1b_lmer)) #qq-plot
qqline(residuals(H1b_lmer)) #line of "perfect normality"
```

## Hypothesis 1 plotting
```{r H1 - plotting, message=FALSE, warning=FALSE}
############
#HYPOTHESIS 1, PLOTTING
############

#Plot log-log with flow z-scores coloring
#First sort participants according to f~d
fss_learning_fdsort <- fss_learning
fss_learning_fdsort$ID <- factor(fss_learning_fdsort$ID, levels=c(1:18))
estimates_slope <- list()
counter = 0
for (x in 1:length(unique(fss_learning_fdsort$ID))) { #iterate through each ID
  estimates_temp <- subset(fss_learning_fdsort, ID==x) #subset one ID at a time
  temp_model <- lm(flow ~ deviation, data=estimates_temp)  #fit model for subsetted ID
  counter = counter + 1 #counter iteration for indexing
  estimates_slope[[counter]] <- coef(summary(temp_model))[2,1] #add values to estimates_slope list
}
estimates_slope <- as.data.frame(rbind(unlist(estimates_slope))) #make list into dataframe
colnames(estimates_slope) <- c(1:18) #rename columns (from V1...V18 to 1...18)
estimates_slope <- sort(estimates_slope, decreasing=F) #sort according to values
part_levels <- as.integer(names(estimates_slope)) #turn into integer values
fss_learning_fdsort$ID <- factor(fss_learning_fdsort$ID, levels=part_levels) #reorder ID factor levels in data according to sorted integers

# plot hypothesis 1a
rq1a <- ggplot(fss_learning_fdsort, aes(ln.cumrun, ln.duration, color = z.flow)) +
  geom_point(alpha=.6, size=1) +
  geom_smooth(method = "lm", se = FALSE, linetype = 1, size = 0.5, color="red") +
  facet_wrap(~ID) +
  scale_color_viridis(name="Z_Flow", guide = guide_colorbar(title.position = "top")) +
  xlab("ln(Cumulative runs)") + ylab("ln(Duration)") +
  labs(title = "A") +
  theme(legend.position = c(x=.8, y=.1),
        legend.background = element_rect(fill="white"),
        legend.box.background = element_rect(colour = "black"),
        legend.direction='horizontal',
        panel.background = element_rect(fill = "white",
                                        colour = "lightgrey",
                                        size = 0.5, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                        colour = "grey"),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="lightgrey", color="black"))

# plot hypothesis 1b
rq1b <- ggplot(fss_learning_fdsort, aes(deviation, flow)) +
  geom_point(alpha=.4, size=1) +
  geom_vline(xintercept=0, linetype=2, alpha=.5) + #line to separate better-than-expected and worse-than-expected performance
  geom_rect(fill="green",alpha=0.005, xmin=min(fss_learning_fdsort$deviation), xmax=0, ymin=0, ymax=max(fss_learning_fdsort$flow)) +
  geom_rect(fill="red",alpha=0.005, xmin=0, xmax=max(fss_learning_fdsort$deviation), ymin=0, ymax=max(fss_learning_fdsort$flow)) +
  geom_smooth(method = "lm", se = FALSE) + 
  xlab("Deviation score") + ylab("Flow score") +
  facet_wrap(~ID) +
  labs(title = "B") +
  theme(legend.position = c(x=.8, y=.1),
        legend.background = element_rect(fill="white"),
        legend.box.background = element_rect(colour = "black"),
        legend.direction='horizontal',
        panel.background = element_rect(fill = "white",
                                        colour = "lightgrey",
                                        size = 0.5, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                        colour = "grey"),
        axis.text.x = element_text(size=8),
        axis.text.y = element_text(size=8),
        axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="lightgrey", color="black"))

#Combine Rq1a and b into panel figure
RQ1 <- ggarrange(rq1a, rq1b, font.label = list(size = 10, color = "black", face = "bold"), common.legend = FALSE, hjust = -0.25, ncol = 2, nrow = 1)

RQ1

ggsave("figure4.pdf", width=12, height=6)
```

# Hypothesis 2
```{r H2 - Effect of experience on the f~d effect, message=FALSE, warning=FALSE}
```

#Hypothesis 2 analyses
```{r H2 - analyses, message=FALSE, warning=FALSE}
############
#HYPOTHESIS 2, ANALYSES: Does the effect of Deviation on Flow (main finding from previous paper) depend on level cumrun (how many runs the participant has played)
############

#Fit model:
H2_lmer <- lmer(flow ~ deviation*cumrun + (deviation*cumrun|participant), data=fss_learning)
summ(H2_lmer)

#Random intercept only
H2_lmer_simple <- lmer(flow ~ deviation*cumrun + (1|participant), data=fss_learning)
summ(H2_lmer_simple)

#Simple slopes:
sim_slopes(H2_lmer_simple, pred = deviation, modx = cumrun, johnson_neyman = TRUE)

#Test model assumptions:
plot(H2_lmer_simple)
plot(resid(H2_lmer_simple), fss_learning$deviation)
plot(resid(H2_lmer_simple), fss_learning$cumrun)
qqnorm(residuals(H2_lmer_simple))
qqline(residuals(H2_lmer_simple))
qqmath(ranef(H2_lmer_simple, condVar = TRUE), strip = TRUE)

#Fit model using robust lmm to confirm result (commented out for increased speed!)
#H2_rlmer_simple <- rlmer(flow ~ deviation*cumrun + (1|participant), data=fss_learning)
#summary(H2_rlmer_simple)
```

## Hypothesis 2, plotting
```{r H2 - plotting, message=FALSE, warning=FALSE}
############
#HYPOTHESIS 2, PLOTTING
############

#Participant-wise visualizations of deviation*cumrun interaction effect
#Calculate slopes and intercepts of deviation at various levels of moderator (cumrun) for each participant.
cumrun_minusSD <- mean(fss_learning$cumrun) - sd(fss_learning$cumrun)
cumrun_meanSD <- mean(fss_learning$cumrun)
cumrun_plusSD <- mean(fss_learning$cumrun) + sd(fss_learning$cumrun)
slope_intercept_dataframes = list()
fss_models <- list()
for (i in 1:length(unique(fss_learning$participant))) {
  temp_model = lm(flow ~ deviation*cumrun, data=subset(fss_learning, ID==i)) #Here we're essentially just fitting LMs to each participant separately.
  flow_interaction <- effect(c("deviation*cumrun"), temp_model, 
                             xlevels=list(deviation=c(-0.2, 0, 0.2), #arbitrary levels, 0 to get intercept
                                          cumrun=c(cumrun_minusSD, cumrun_meanSD, cumrun_plusSD), #values are -1SD and +1SD
                                          se=TRUE, confidence.level=.95, typical=mean))
  flow_interaction <- as.data.frame(flow_interaction)
  flow_interaction$ID <- as.factor(i)
  flow_interaction$cumrun <- as.factor(flow_interaction$cumrun)
  #flow_interaction$ID <- as.factor(flow_interaction$ID)
  flow_interaction <- flow_interaction %>% mutate(slope_minusSD = fit[3]-fit[1], #slope is the difference in FIT between 1 unit increment in deviation
                                                  slope_MEAN = fit[6]-fit[4],
                                                  slope_plusSD = fit[9]-fit[7],
                                                  intercept_minusSD = fit[2], #intercept is the value of fit when deviation = 0
                                                  intercept_MEAN = fit[5],
                                                  intercept_plusSD = fit[8])
  slope_intercept_dataframes = rbind(slope_intercept_dataframes, flow_interaction)
  fss_models[[i]] = temp_model
}

#Grab p-values, sort from lowest to highest, and adjust using Bonferroni-Holm
fss_Pvalues <- list()
for (i in 1:18) {
  fss_Pvalues[[i]] <- anova(fss_models[[i]])$"Pr(>F)"[3]
}
fss_Pvalues <- as.data.frame(fss_Pvalues) 
colnames(fss_Pvalues) <- c(1:18)
fss_Pvalues <- sort(fss_Pvalues)
fss_Pvalues_holm <- p.adjust(fss_Pvalues, "holm")

#Grab order of participants for plotting
part_levels <- as.integer(names(fss_Pvalues))

#Reorder levels in data; now according to p-value, from low (significant) to high (nonsignificant):
fss_learning$ID <- factor(fss_learning$ID, levels=part_levels) #NOTE 10.1. this causes issue when subsetting using ">" on ID

#Plot
flow_fig6x3 <- ggplot(fss_learning, aes(deviation, flow)) +
  geom_point(alpha=.2, size=1.5) +
  #geom_smooth(method = "lm", se = FALSE, fullrange=TRUE) + 
  facet_wrap(~ID, nrow=3, ncol=6) + 
  geom_abline(data = slope_intercept_dataframes, aes(intercept=intercept_minusSD, slope=slope_minusSD, linetype="-1 SD\n(8.81 runs)"), size=0.8) + #-1SD is at 8.81 runs
  geom_abline(data = slope_intercept_dataframes, aes(intercept=intercept_plusSD, slope=slope_plusSD, linetype="+1 SD\n(32.19 runs)"), size=0.8) +  #+1SD is at 32.19 runs
  geom_vline(xintercept=0, linetype=2, alpha=.5) + #line to separate better-than-expected and worse-than-expected performance
  geom_rect(fill="green",alpha=0.005, xmin=min(fss_learning$deviation), xmax=0, ymin=0, ymax=max(fss_learning$flow)) +
  geom_rect(fill="red",alpha=0.005, xmin=0, xmax=max(fss_learning$deviation), ymin=0, ymax=max(fss_learning$flow)) +
  labs(colour = NULL, linetype = "Cumulative runs", title = "A") + xlab("Deviation score") + ylab("Flow score") +
  #scale_x_discrete(breaks = c()) +
  theme_bw(base_size=13) + 
  theme(legend.justification = "top",
        legend.background = element_rect(fill="white"),
        legend.box.background = element_rect(colour = "black"),
        legend.text = element_text(size=8),
        legend.title = element_text(size=9),
        axis.text.x = element_text(size=9),
        panel.background = element_rect(fill = "white",
                                        colour = "lightgrey",
                                        size = 0.5, linetype = "solid"),
        panel.border = element_blank(), 
        panel.grid.major = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="lightgray"))

#Participant-wise Sliding-window regression, custom function
sliding_window_within <- function(window_width = 10, result=2) {
  estimates_IDwise <- list()
  counter = 0
  for (x in 1:length(unique(fss_learning$participant))) { #iterate through all participants
    for (i in 1:(length(unique(fss_learning$cumrun))+1-window_width)) {
      fss_learning_temp <- subset(fss_learning, ID==x & cumrun >= i & cumrun < i+window_width) #subset based on selected window width
      
      ###calculate deviation score again for each segment
      fss_learning_temp <- fss_learning_temp %>% group_by(participant) %>%
        mutate(learning_curve = predict(lm(ln.duration ~ ln.cumrun)), deviation = ln.duration - learning_curve)
      ###
      temp_model <- lm(flow ~ deviation, data=fss_learning_temp)
      counter = counter + 1
      estimates_IDwise[[counter]] <- coef(summary(temp_model))[2,1]
    }
  }
  
  estimates_IDwise <- as.data.frame(unlist(estimates_IDwise)) 
  estimates_IDwise <- estimates_IDwise %>% mutate(sliding_window=rep(c(1:(length(unique(fss_learning$cumrun))+1-window_width)), 18),
                                                  ID=rep(1:18, each=length(unique(fss_learning$cumrun))+1-window_width)) #18 is number of participants
 
  names(estimates_IDwise)[1] <- "estimate"
  permuted_estimates <- estimates_IDwise %>% group_by(ID) %>% 
    transform(estimate = sample(estimate))
  
  #code for sorting facets based on slope BEGINS HERE
  #First, fit linear model separately for each participant with sliding window as predictor and slope estimate as DV
  #then grab the slope estimate of that model, and sort according to the values (this is essentially the same as above for p-values)
  estimates_slope <- list()
  counter = 0
  for (x in 1:length(unique(estimates_IDwise$ID))) {
    estimates_temp <- subset(estimates_IDwise, ID==x)
    temp_model <- lm(estimate ~ sliding_window, data=estimates_temp) 
    counter = counter + 1
    estimates_slope[[counter]] <- coef(summary(temp_model))[2,1]

  }
  
  estimates_slope <- as.data.frame(rbind(unlist(estimates_slope)))
  colnames(estimates_slope) <- c(1:18)
  estimates_slope <- sort(estimates_slope)
  part_levels <- as.integer(names(estimates_slope))
  estimates_IDwise$ID <- factor(estimates_IDwise$ID, levels=part_levels)
  #code for sorting facets based on slope ENDS HERE
  
  #choose which result to return (1=Participant-wise figure, 2=DATA VECTOR: permuted estimates, 3=DATA VECTOR: observed estimates):
  #participant-wise
  if (result == 1) {
    result <- ggplot(estimates_IDwise, aes(x=sliding_window, y=estimate)) + geom_line() + 
      ylab("Slope estimate") + xlab(NULL) + #xlab should be "sliding window" but changed now for cogcarsim #2 paper
      theme_bw(base_size=13) + scale_x_continuous(breaks = c(1, 10, 20, 31)) +
      theme(axis.text.x = element_text(size=8)) +
      facet_wrap("ID", ncol=6, nrow=3) + geom_smooth(method="lm", se=FALSE, size=0.3) + labs(title = "B") #note that title and ncol, nrow for facet_wrap are now specific to cogcarsim #2 paper!
  }
  
  #permuted estimates:
  else if (result == 2) {
    result <- permuted_estimates
  }
  
  #observed estimates:
  else {
    result <- estimates_IDwise
  }
  return(result)
}

#Combine the simple slopes figure (flow_fig6x3) with sliding windows regression figure.
#Use Minimum Width Envelope of the actual observations (sliding windows of slopes within participants)
#MWE of actual observations
estimates_IDwise_wide <- sliding_window_within(10,3) %>% pivot_wider(names_from = sliding_window, values_from = estimate) %>% dplyr::select(-ID)
curve <- findcurves(estimates_IDwise_wide)
idx = c(1:31)
rng <- range(curve[,c("mean0","lo","up")])

mwe_fig <- ggplot(cbind(curve, idx), aes(x = idx, y = mean0)) +
  geom_ribbon(aes(ymin = lo, ymax = up, fill = "red"), alpha = 0.2, linetype = 0) +
  geom_ribbon(aes(ymin = lo0, ymax = up0, fill = "blue"), alpha = 0.2, linetype = 0) +
  ylim(-21.5,0) + scale_x_continuous(breaks=c(1, 10, 20, 31)) + #change ylim if ribbon seems to be buggy
  geom_line(size = 0.75) +
  labs(title = "", x = NULL, y = NULL) + #title = "" to force empty space near top
  theme_bw() +
  theme(legend.position = "none")

combined_plot1 <- ggarrange(sliding_window_within(10,1), mwe_fig, ncol=2, nrow=1, widths = c(2,1.5))
combined_plot1 <- annotate_figure(combined_plot1, 
                                  bottom = text_grob("Sliding window", color = "black", size=14),
                                  fig.lab = NULL, fig.lab.face = NULL)

ggarrange(flow_fig6x3, combined_plot1, ncol=1, nrow=2)

ggsave("combined_interaction_sliding.pdf", width=10, height=12)


```

# Other plots
```{r Other plots, message=FALSE, warning=FALSE}

############
#Plot basic stats
############

# Basic violin plot, flow
flow1 <- fss_items %>%
  mutate(participant = fct_reorder(participant, desc(flow), .fun='median')) %>%
  ggplot( aes(x=participant, y=flow)) + 
  geom_violin() + geom_hline(aes(yintercept=median(flow)), color="firebrick") +
  geom_jitter(color="black", size=0.6, alpha=0.2, width=0.07) +
  # geom_half_violin(trim=FALSE, fill="gray", side = "r") +
  # geom_boxplot(width=0.1, outlier.shape = NA) +
  geom_hline(yintercept = median(fss_items$flow), color = "firebrick") +
  #geom_text(aes(length(unique(fss_items$participant)), median(fss_items$flow), label = "med.\nFlow", vjust = -1)) +
  labs(title="A", x="Participant", y = "Flow") +
  ylim(2, 7) +
  stat_summary(fun = median, geom = "point",
               size = 2, color="firebrick") +
  theme_classic(base_size=14)

#Session-wise flow scores, new data
flow2 <- fss_learning %>% 
  mutate(session = fct_reorder(as.factor(session), desc(flow), .fun='median')) %>% 
  ggplot(aes(as.factor(session), flow)) + 
  geom_violin() + geom_hline(aes(yintercept=median(flow)), color="firebrick") +
  geom_jitter(color="black", size=0.6, alpha=0.2, width=0.07) +
  ylab(NULL) + xlab("Session") +
  ylim(2, 7) +
  labs(title="B") +
  stat_summary(fun = median, geom = "point",
               size = 2.5, color="firebrick") +
  theme_classic(base_size=14)

#Combine flow1 and flow2 figures
ggarrange(flow1, flow2, font.label = list(size = 10, color = "black", face = "bold"), common.legend = FALSE, hjust = -0.25, ncol = 2, nrow = 1)

ggsave("figure3_alternative.pdf", width=12, height=4) #twiikkaa kokoa, ei lopullinen

############
#Visualize effect of deviation vs. skill
############

testilmer <- lmer(flow ~ ln.duration + deviation + (1|ID), data=fss_learning)

nd <- with(fss_learning,
           rbind(data.frame(expand.grid(ID=levels(ID),
                                        ln.duration=seq(min(ln.duration),max(ln.duration),length=51)),
                            deviation=mean(deviation),focus="ln.duration"), #to predict for values of duration while holding deviation constant at its mean
                 data.frame(expand.grid(ID=levels(ID),
                                        deviation=seq(min(deviation),max(deviation),length=51)),
                            ln.duration=mean(ln.duration),focus="deviation"))) #to predict for values of deviation while holding duration constant at its mean

nd$val <- with(nd, c(ln.duration[focus=="ln.duration"], deviation[focus=="deviation"]))
nd$ID <- factor(nd$ID, levels=1:18)
nd$focus <- factor(nd$focus, labels = c("log(Duration)", "Deviation score"))

pframe <- data.frame(nd,resp=predict(testilmer,newdata=nd))

ggplot(pframe,aes(x=val,y=resp,colour=ID)) + geom_line()+
  facet_wrap(~focus,scale="free_x") + xlab("Value") + ylab("Flow score") + theme_bw(base_size=12) +
  ylim(2.5, 6.5) + scale_colour_discrete(name = "Participant") + theme(legend.text = element_text(size=9),
                                                                       legend.title = element_text(size=12))

ggsave("partial_effects_lmer.pdf", width=6, height=5)


```

# Appendix
```{r Appendix, message=FALSE, warning=FALSE}
```

## Background variables
```{r Background variables, message=FALSE, warning=FALSE}
############
#Background variables
############

names(background19)[2] <- "gender"
names(background19)[3] <- "age"
names(background19)[4] <- "license"
names(background19)[6] <- "driving_experience"
names(background19)[7] <- "gaming_experience" #unclear how this is coded in 2017 so not analysed yet!
background19 <- background19 %>% dplyr::select(ID, gender, age, driving_experience, gaming_experience)
background19$gender <- factor(background19$gender, labels=c("m", "f"))
background19$driving_experience <- as.numeric(as.character((factor(background19$driving_experience, labels=c("2", "3", "5", "4", "1")))))
background19$gaming_experience <- as.numeric(as.character((factor(background19$gaming_experience, labels=c("3", "4", "1", "2")))))

background17 <- data.frame(ID = seq(1:9),
                           gender = c("m", "m", "f", "f", "m", "m", "f", "m", "m"),
                           age = c(27,23,23,28,27,22,31,38,25), 
                           driving_experience = c(2,3,1,1,4,4,3,5,3),
                           gaming_experience = c(4,4,3,4,4,3,1,3,4))

background_flowcar <- rbind(background17, background19)
background_flowcar <- background_flowcar %>% mutate(ID = seq(1:18))

#expand each row into 40, except for participant 8 (who has 39 rows)
background_flowcar <- background_flowcar[rep(row.names(background_flowcar), each=40), 1:5]

#remove one row, can be any row, of participant 8 (row number 306; session 6, run 1, is missing in fss_learning; here doesn't matter which row since background is measured once per participant)
background_flowcar <- background_flowcar[-306,] %>% dplyr::select(-ID)
fss_learning_full <- cbind(fss_learning, background_flowcar)

#Driving experience coding:
# 1 = 0-1000
# 2 = 1000-10000
# 3 = 10000-30000
# 4 = 30000-100000
# 5 = 100000-300000

#Gaming experience coding:
# 1 = Little / none
# 2 = A few times per year
# 3 = 1-3 hours per month
# 4 = At least one hour per week

apx1 <- ggplot(fss_learning_full, aes(cumrun, duration, colour=gender)) +
  geom_point(alpha=.1, size=2) +
  geom_smooth(method = "nls", se = FALSE, formula = 'y~a*x^b') +
  theme_bw(base_size = 14) +
  theme(legend.position = c(x=.7, y=.7),
        legend.background = element_rect(fill="white"),
        legend.text = element_text(size=14),
        panel.border = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="grey")) +
  ylab("Run duration in seconds") + xlab(NULL) +
  labs(title=NULL) +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  scale_colour_manual(name=NULL, labels=c("Female", "Male"), values=c("red3", "blue3"))

apx2 <- ggplot(fss_learning_full, aes(cumrun, duration, colour=factor(driving_experience))) +
  geom_point(alpha=.1, size=2) +
  geom_smooth(method = "nls", se = FALSE, formula = 'y~a*x^b') +
  theme_bw(base_size = 14) +
  theme(legend.position = c(x=.6, y=.7),
        legend.background = element_rect(fill="white"),
        #legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        panel.border = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="grey")) +
  ylab(NULL) + xlab(NULL) +
  labs(title=NULL) +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  scale_colour_discrete(name=NULL, labels=c("No lifetime driving experience", "1000-10000 km", "10000-30000 km", "30000-100000 km", "100000-300000 km"))

apx3 <- ggplot(fss_learning_full, aes(cumrun, duration, colour=factor(gaming_experience))) +
  geom_point(alpha=.1, size=2) +
  geom_smooth(method = "nls", se = FALSE, formula = 'y~a*x^b') +
  theme_bw(base_size = 14) +
  theme(legend.position = c(x=.6, y=.7),
        legend.background = element_rect(fill="white"),
        #legend.title = element_text(size=12),
        legend.text = element_text(size=10),
        panel.border = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="grey")) +
  ylab(NULL) + xlab(NULL) +
  labs(title=NULL) +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  scale_colour_discrete(name=NULL, labels=c("Little/no gaming experience", "A few times per year", "1-3 hours per month", "At least one hour per week"))

# combine apx1-3 (gender, driving- and gaming experience)
background_apx <- ggarrange(apx1, apx2, apx3, font.label = list(size = 10, color = "black", face = "bold"), 
                            common.legend = FALSE, hjust = -0.25, ncol = 3, nrow = 1)

background_apx <- annotate_figure(background_apx, 
                                  bottom = text_grob("Cumulative runs", color = "black", size=14),
                                  fig.lab = NULL, fig.lab.face = NULL)

background_apx

ggsave("background_apx.pdf", width=12, height=4)


# Basic violin plot, perceived importance items
apx4 <- fss_items %>%
  mutate(participant = fct_reorder(participant, desc(pi_total), .fun='median')) %>%
  ggplot( aes(x=participant, y=pi_total)) + 
  geom_violin() + geom_hline(aes(yintercept=median(pi_total)), color="firebrick") +
  geom_jitter(color="black", size=0.6, alpha=0.2, width=0.07) +
  # geom_half_violin(trim=FALSE, fill="gray", side = "r") +
  # geom_boxplot(width=0.1, outlier.shape = NA) +
  geom_hline(yintercept = median(fss_items$pi_total), color = "firebrick") +
  #geom_text(aes(length(unique(fss_items$participant)), median(fss_items$pi_total), label = "med.\nFlow", vjust = -1)) +
  labs(title="Perceived importance", x=NULL, y = NULL) +
  ylim(2, 7) +
  stat_summary(fun = median, geom = "point",
               size = 2, color="firebrick") +
  theme_classic(base_size=12)

apx5 <- fss_items %>%
  mutate(participant = fct_reorder(participant, desc(fluency), .fun='median')) %>%
  ggplot( aes(x=participant, y=fluency)) + 
  geom_violin() + geom_hline(aes(yintercept=median(fluency)), color="firebrick") +
  geom_jitter(color="black", size=0.6, alpha=0.2, width=0.07) +
  # geom_half_violin(trim=FALSE, fill="gray", side = "r") +
  # geom_boxplot(width=0.1, outlier.shape = NA) +
  geom_hline(yintercept = median(fss_items$fluency), color = "firebrick") +
  #geom_text(aes(length(unique(fss_items$participant)), median(fss_items$fluency), label = "med.\nFlow", vjust = -1)) +
  labs(title="Fluency of performance", x=NULL, y = NULL) +
  ylim(2, 7) +
  stat_summary(fun = median, geom = "point",
               size = 2, color="firebrick") +
  theme_classic(base_size=12)

apx6 <- fss_items %>%
  mutate(participant = fct_reorder(participant, desc(absorption), .fun='median')) %>%
  ggplot( aes(x=participant, y=absorption)) + 
  geom_violin() + geom_hline(aes(yintercept=median(absorption)), color="firebrick") +
  geom_jitter(color="black", size=0.6, alpha=0.2, width=0.07) +
  # geom_half_violin(trim=FALSE, fill="gray", side = "r") +
  # geom_boxplot(width=0.1, outlier.shape = NA) +
  geom_hline(yintercept = median(fss_items$absorption), color = "firebrick") +
  #geom_text(aes(length(unique(fss_items$participant)), median(fss_items$absorption), label = "med.\nFlow", vjust = -1)) +
  labs(title="Absorption by activity", x=NULL, y = NULL) +
  ylim(2, 7) +
  stat_summary(fun = median, geom = "point",
               size = 2, color="firebrick") +
  theme_classic(base_size=12)

# combine apx4-6 (PI items, fluency and absorption)
scales_apx <- ggarrange(apx4, apx5, apx6, font.label = list(size = 10, color = "black", face = "bold"), 
                        common.legend = FALSE, hjust = -0.25, ncol = 3, nrow = 1)

scales_apx <- annotate_figure(scales_apx, 
                              bottom = text_grob("Participant", color = "black", size=14),
                              fig.lab = NULL, fig.lab.face = NULL)

scales_apx

ggsave("scales_apx.pdf", width=12, height=4)
```

## Lagged flow regressions
```{r Lagged flow regressions, message=FALSE, warning=FALSE}
############
#Lagged flow regressions
############

#NOTE: dplyr::mutate works, but plyr:mutate() doesn't work correctly with group_by and lag()
lag_fss_learning <- fss_learning %>% group_by(participant) %>% dplyr::mutate(lag_flow1 = lag(flow, 1), lag_flow2 = lag(flow, 2), lag_flow3 = lag(flow, 3), 
                                                                      lag_flow4 = lag(flow, 4), lag_flow5 = lag(flow, 5),
                                                                      lag_deviation1 = lag(deviation, 1), lag_deviation2 = lag(deviation, 2), lag_deviation3 = lag(deviation, 3),
                                                                      lag_deviation4 = lag(deviation, 4), lag_deviation5 = lag(deviation, 5),
                                                                      lag_duration1 = lag(ln.duration, 1), lag_duration2 = lag(ln.duration, 2),
                                                                      lag_duration3 = lag(ln.duration, 3), lag_duration4 = lag(ln.duration, 4),
                                                                      duration_dif = lag(ln.duration, 1) - ln.duration)

lag_fss_learning_longer <- lag_fss_learning %>% dplyr::select(ID, flow, lag_flow1, lag_flow2, lag_flow3, lag_flow4) %>%
  gather(key=lag_amount, value=lagged_flow_score, lag_flow1, lag_flow2, lag_flow3, lag_flow4)

# Plotting
ggplot(lag_fss_learning_longer, aes(lagged_flow_score, flow, color=lag_amount, size=lag_amount)) + 
  geom_smooth(method = "lm", fullrange=T, se=F) + 
  geom_point(alpha=.15) +
  facet_wrap("participant") +
  scale_colour_viridis_d(labels=c("Flow lag = 1", "Flow lag = 2", "Flow lag = 3", "Flow lag = 4"),
                         direction = -1) +
  scale_size_discrete(range=c(1, 0.2), labels=c("Flow lag = 1", "Flow lag = 2", "Flow lag = 3", "Flow lag = 4")) +
  theme_bw(base_size=13) + xlab("Lagged flow score") + ylab("Flow score") +
  guides(size=guide_legend(title=NULL), color=guide_legend(title=NULL)) +
  theme(legend.position = c(x=.8, y=.1),
        legend.background = element_rect(fill="white"),
        axis.text.x = element_text(size=8),
        panel.border = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="grey"))

ggsave("lagged_flow.pdf", width=7, height=6)
```

## Multilevel mediation analysis
```{r Multilevel mediation analysis, message=FALSE, warning=FALSE}
############
#Multilevel Mediation Analysis
############

# Multilevel models assume that the observed variables have at least two potential levels of variation. 
# Because temporal lag was experimentally manipulated within subjects, it does not vary between subjects. 
# On the other hand, hit rates (HR) vary both between and within participants: At the lower (within-person) level, 
# HR varies from trial to trial. At the upper level, we may also expect that HR varies, on average, 
# between participants. We are most interested in the within-person process, and therefore it is useful to 
# transform the variables such that these two levels are explicitly separated from each other (Bolger & Laurenceau, 2013). 
# Notice that this transformation is not strictly required, but this reasoning suggests that it is often useful and meaningful in data sets where the predictor values vary both between and within subjects. 
# We first averaged the grand-mean-centered trial-level HR for each person to create a between-person component of HR. 
# We subtracted these means from the raw HR to create within-subject trial-by-trial deviations from the subject-means that 
# represent an entirely within-person version of HR. Isolating the within-person process from variables can 
# be done by using bmlm’s isolate() function:

#using mlm() from https://link.springer.com/article/10.3758/s13428-017-0980-9#Tab1
#https://mvuorre.github.io/bmlm/articles/bmlm-blch9/bmlm-blch9.html
#need first to remove "NA" value from duration_dif (first value)

fss2 <- fss %>% mutate(session = rep(rep(1:8, each = 5), 18))
fss2 <- fss2[-which(fss2$session==6 & fss2$run==1 & fss2$participant==8),]

lag_fss_learning2 <- lag_fss_learning %>% dplyr::select(duration_dif, deviation, flow)
lag_fss_learning2$duration_dif[is.na(lag_fss_learning2$duration_dif)] <- 0
lag_fss_learning2 <- isolate(lag_fss_learning2, by = "participant",
                             value = c("duration_dif", "deviation", "flow")) #separate between subjects variability
lag_fss_learning2 <- as.data.frame(lag_fss_learning2) #change into dataframe -- cannot be tibble for mlm()!

fit <- mlm(d = lag_fss_learning2, 
           id = "participant",
           x = "duration_dif_cw",
           m = "deviation_cw",
           y = "flow_cw",
           iter = 1000, 
           cores = 1)

mlm_summary(fit)

mlm_path_plot(fit, level = .95, text = T,
              xlab = "Feedback",
              mlab = "Deviation score",
              ylab = "Self-reported flow", digits = 2)

mlm_pars_plot(fit, pars = "u_me", type = "coef", level = .90) #mediated effect for all participants, 90% credible intervals



```

## Scatterplot matrix of game variables
```{r Scatterplot matrix of game variables, message=FALSE, warning=FALSE}
############
#Scatterplot matrix of game variables:
############

game_data %>%
  dplyr::select(cumrun, min_velocity, max_velocity, end_velocity, avg_velocity, collisions, speed_drops, duration) %>% 
  pairs.panels(method = "spearman", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             smoother = TRUE) # show correlation ellipses

```

## Cronbach's alpha for FSS
```{r Alpha for FSS, message=FALSE, warning=FALSE}
############
#Custom Cronbach's alpha calculator for each trial. 1= full scale, 2 = fluency, 3 = absorption
############

alpha_calculator <- function(subscale=1) {
  cronbachs_alphas <- cbind(fss2, fss_learning_full)
  #cronbachs_alphas <- cbind(fss2[-306,], fss_learning_full) #remove one row from participant 8 (session 6 run 1), this row is missing in fss_learning (game data). NOT NECESSARY, WAS DONE EARLIER!
  #coud also use: fss2[-which(fss2$session==6 & fss2$run==1 & fss2$participant==8),] to double check row is correct
  cronbachs_alphas$run_session <- as.integer(interaction(fss_learning$run, fss_learning$session)) #expand run integer into 1:40
  alphas_list <- list()
  alphas <- list()
  counter = 0
  if (subscale == 1) {
    for (i in 1:length(unique(cronbachs_alphas$run_session))) { #calculate cronbach's alpha separately for each run
      counter = counter + 1
      alphas_list[[counter]] <- subset(cronbachs_alphas, run_session==i) %>% dplyr::select(Q2, Q4, Q5, Q6, Q7, Q8, Q9, Q10) %>% psych::alpha()
      alphas[[counter]] <- summary(alphas_list[[counter]])$std.alpha
    }
    alphas <- as.data.frame(unlist(alphas))
    names(alphas)[1] <- "alpha"
    
    return(ggplot(alphas, aes(x=alpha)) + geom_histogram(alpha=.7, color="grey", fill="firebrick") +
            scale_x_continuous(name = NULL, breaks=c(0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95)) +
            ylab("Count") + labs(title="Full FSS scale") + scale_y_continuous(breaks=c(1,2,3,4,5,6,7,8,9), limits=c(0,9)) +
            #coord_cartesian(xlim = c(0.55, max(alphas))) +
            theme_classic(base_size=12))

  }
  else if (subscale == 2) {
    for (i in 1:length(unique(cronbachs_alphas$run_session))) { #calculate cronbach's alpha separately for each run
      counter = counter + 1
      alphas_list[[counter]] <- subset(cronbachs_alphas, run_session==i) %>% dplyr::select(Q2, Q4, Q5, Q7, Q8, Q9) %>% psych::alpha()
      alphas[[counter]] <- summary(alphas_list[[counter]])$std.alpha
    }
    alphas <- as.data.frame(unlist(alphas))
    names(alphas)[1] <- "alpha"
    
    return(ggplot(alphas, aes(x=alpha)) + geom_histogram(alpha=.7, color="grey", fill="firebrick") +
             scale_x_continuous(name = NULL, breaks=c(0.55, 0.65, 0.75, 0.85, 0.95)) +
             ylab(NULL) + labs(title="Fluency of performance") + scale_y_continuous(breaks=c(1,2,3,4,5,6,7,8,9), limits=c(0,9)) +
             #coord_cartesian(xlim = c(0.55, max(alphas))) +
             theme_classic(base_size=12))
  }
  else if (subscale == 3) {
    for (i in 1:length(unique(cronbachs_alphas$run_session))) { #calculate cronbach's alpha separately for each run
      counter = counter + 1
      alphas_list[[counter]] <- subset(cronbachs_alphas, run_session==i) %>% dplyr::select(Q6, Q10) %>% psych::alpha() 
      alphas[[counter]] <- summary(alphas_list[[counter]])$std.alpha
    }
    alphas <- as.data.frame(unlist(alphas))
    names(alphas)[1] <- "alpha"
    
    return(ggplot(alphas, aes(x=alpha)) + geom_histogram(alpha=.7, color="grey", fill="firebrick") +
             scale_x_continuous(name = NULL, breaks=c(0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1)) +
             ylab(NULL) + labs(title="Absorption by activity") + scale_y_continuous(breaks=c(1,2,3,4,5,6,7,8,9), limits=c(0,9)) +
             #coord_cartesian(xlim = c(0.55, max(alphas))) +
             theme_classic(base_size=12))
  }
}

#all in the same figure
alpha_reliability <- ggarrange(alpha_calculator(1), alpha_calculator(2), alpha_calculator(3), 
          font.label = list(size = 10, color = "black", face = "bold"), 
          common.legend = TRUE, legend="bottom", hjust = -0.25, ncol = 3, nrow = 1)

alpha_reliability <- annotate_figure(alpha_reliability, 
                                     bottom = text_grob("Cronbach's alpha", color = "black", size=14),
                                     fig.lab = NULL, fig.lab.face = NULL)

alpha_reliability

ggsave("histogram_alpha.pdf", width=10, height=5) #uncomment to save figure as pdf

```

## Spread of measurement sessions
```{r Spread of measurement sessions, message=FALSE, warning=FALSE}

############
#Visualize spread of measurement sessions in time
############

game_data2 <- game_data %>% group_by(participant, session) %>% dplyr::summarize(session = mean(session), date=mean.Date(as.Date(date))) %>%
  mutate(year=as.factor(ifelse(as.integer(participant)>9, 2019, 2017)))

#Tricky way to make the plotting work: here we substitute year 2017 with year 2019 so that free-scale faceting works better
game_data2$date <- sub("2017", "2019", game_data2$date)
game_data2$date <- as.Date(game_data2$date)

game_data2 <- game_data2 %>% mutate(physiology = as.factor(ifelse(session == 2 | session == 3 | session == 4, 0, 1)))

#Create separate dataframes to call for geom_text (enables drawing two different texts for facetet grids)
data.year2017 <- data.frame(participant=6,date=as.Date("2020-1-6"),session=NA,year=as.factor(2017))
data.year2019 <- data.frame(participant=6,date=as.Date("2020-1-6"),session=NA,year=as.factor(2019))

#Add gap between measurement dates
game_data2 <- game_data2 %>% group_by(participant) %>% dplyr::mutate(gap = as.integer(date - lag(date, 1)), mean_gap = mean(gap, na.rm=TRUE)) #note that with lag(), must use dplyr::mutate()

ggplot(game_data2, aes(participant, date, group=factor(session))) + geom_point(aes(color=physiology), size=2.5, alpha=.8) + coord_flip() + 
  facet_wrap("year") + #use scales="free_x" if year 2017 is NOT replaced by 2019
  scale_y_date(date_breaks = "weeks" , date_labels = "%b-%d") +
  scale_color_discrete(name=NULL, labels=c("No physiologial measures", "Physiological measures")) +
  theme(axis.text.x = element_text(angle = -60, hjust = 0),
        axis.text.y = element_blank(), #hide participant numbers for anonymity
        panel.background = element_rect(fill = "white",
                                        colour = "lightgrey",
                                        size = 0.5),
        legend.key=element_blank(),
        legend.position="bottom",
        axis.title.y = element_text(size=14),
        axis.title.x = element_text(size=14),
        legend.text = element_text(size=12),
        panel.border = element_blank(),
        panel.grid.major = element_line(size = 0.25, colour = "grey"),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="lightgray"),
        strip.text.x = element_text(size=12)) + 
  geom_hline(yintercept = as.Date("2020-01-1"), linetype="dashed", color ="blue", size=0.5) +
  geom_text(aes(label=session), hjust=0, vjust=0, size=3) +
  geom_text(data=data.year2017, aes(x=participant, y=date), inherit.aes=FALSE, label="2018", 
            color="blue", size=3) +
  geom_text(data=data.year2019, aes(x=participant, y=date), inherit.aes=FALSE, label="2020", 
            color="blue", size=3) +
  ylab(NULL) + xlab("Participants")

ggsave("measurement_dates.pdf", width=10, height=5) #uncomment to save figure as pdf

```

## Effect of perceived competence
```{r Effect of perceived competence, message=FALSE, warning=FALSE}
############
#Effect of perceived competence
############

#Perceived competence increases session by session for participants 2-6, 10, 13-14, 16-17, i.e. 10/18, 55% of the participants
#Overall trend and effect is positive.
fss_learning_tempo <- fss_learning %>% dplyr::select(session, ID, comp2) %>% na.omit()
fss_learning_tempo$ID <- factor(fss_learning_tempo$ID, levels=c(1:18))

#Sort participants according to slope values
estimates_slope <- list()
counter = 0
for (x in 1:length(unique(fss_learning_tempo$ID))) { #iterate through each ID
  estimates_temp <- subset(fss_learning_tempo, ID==x) #subset one ID at a time
  temp_model <- lm(comp2 ~ session, data=estimates_temp)  #fit model for subsetted ID
  counter = counter + 1 #counter iteration for indexing
  estimates_slope[[counter]] <- coef(summary(temp_model))[2,1] #add values to estimates_slope list
}
estimates_slope <- as.data.frame(rbind(unlist(estimates_slope))) #make list into dataframe
colnames(estimates_slope) <- c(1:18) #rename columns (from V1...V18 to 1...18)
estimates_slope <- sort(estimates_slope, decreasing=T) #sort according to values
part_levels <- as.integer(names(estimates_slope)) #turn into integer values
fss_learning_tempo$ID <- factor(fss_learning_tempo$ID, levels=part_levels) #reorder ID factor levels in data according to sorted integers

perc_comp_lmer1 <- lmer(comp2 ~ session + (1|ID), data=fss_learning_tempo)
perc_comp_lmer2 <- lmer(comp2 ~ session + (session|ID), data=fss_learning_tempo)

ggplot(fss_learning_tempo, aes(session, comp2)) + geom_point(alpha=.45) +
  geom_line(aes(y=predict(perc_comp_lmer2), color="Random slope and intercept"), size=0.6) +
  geom_line(aes(y=predict(perc_comp_lmer1), color="Random intercept"), size=0.6) +
  labs(color=NULL) +
  facet_wrap("ID") + ylab("Perceived competence") + xlab("Session") + theme_bw() +
  theme(legend.position = c(x=.8, y=.1),
        legend.background = element_rect(fill="white"),
        legend.box.background = element_rect(colour = "black"),
        legend.text = element_text(size=10),
        axis.text.x = element_text(size=9),
        panel.background = element_rect(fill = "white",
                                        colour = "lightgrey",
                                        size = 0.5, linetype = "solid"),
        panel.border = element_blank(), 
        panel.grid.major = element_line(size = 0.25, linetype = 'solid', colour = "grey"),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
        strip.background=element_rect(fill="lightgray"))
  
ggsave("perceived_importance_session.pdf", width=7.5, height=6) #uncomment to save figure as pdf
```

## Participant-wise partial regression plots
```{r Participant-wise partial regression plots, message=FALSE, warning=FALSE}

############
#Participant-wise partial regression plots
############

###Effect of duration on flow controlling for deviation##
skill_deviation <- fss_learning %>% dplyr::select(session, ID, deviation, ln.duration, flow) %>% na.omit()
skill_deviation$ID <- factor(skill_deviation$ID, levels=c(1:18))

#Remove effect of deviation on flow
residuals_list <- list()
counter = 0
for (x in 1:length(unique(skill_deviation$ID))) { #iterate through each ID
  residuals_temp <- subset(skill_deviation, ID==x) #subset one ID at a time
  residuals_temp <- resid(lm(flow ~ deviation, data=residuals_temp))  #remove effect of deviation on flow
  counter = counter + 1 #counter iteration for indexing
  residuals_list[[counter]] <- residuals_temp #add values to residuals_list list
}

residuals_list <- as.data.frame(unlist(residuals_list)) #make list into dataframe
names(residuals_list)[1] <- "resid_flow_dev"

#Remove effect of deviation on ln.duration
residuals_list2 <- list()
counter = 0
for (x in 1:length(unique(skill_deviation$ID))) { #iterate through each ID
  residuals_temp <- subset(skill_deviation, ID==x) #subset one ID at a time
  residuals_temp <- resid(lm(ln.duration ~ deviation, data=residuals_temp))  #remove effect of deviation on flow
  counter = counter + 1 #counter iteration for indexing
  residuals_list2[[counter]] <- residuals_temp #add values to residuals_list list
}

residuals_list2 <- as.data.frame(unlist(residuals_list2)) #make list into dataframe
names(residuals_list2)[1] <- "resid_dur_dev"


###Effect of deviation on flow controlling for duration###
residuals_list3 <- list()
counter = 0
for (x in 1:length(unique(skill_deviation$ID))) { #iterate through each ID
  residuals_temp <- subset(skill_deviation, ID==x) #subset one ID at a time
  residuals_temp <- resid(lm(flow ~ ln.duration, data=residuals_temp))  #remove effect of ln.duration on flow
  counter = counter + 1 #counter iteration for indexing
  residuals_list3[[counter]] <- residuals_temp #add values to residuals_list list
}

residuals_list3 <- as.data.frame(unlist(residuals_list3)) #make list into dataframe
names(residuals_list3)[1] <- "resid_flow_dur"

#Remove effect of deviation on ln.duration
residuals_list4 <- list()
counter = 0
for (x in 1:length(unique(skill_deviation$ID))) { #iterate through each ID
  residuals_temp <- subset(skill_deviation, ID==x) #subset one ID at a time
  residuals_temp <- resid(lm(deviation ~ ln.duration, data=residuals_temp))  #remove effect of deviation on flow
  counter = counter + 1 #counter iteration for indexing
  residuals_list4[[counter]] <- residuals_temp #add values to residuals_list list
}

residuals_list4 <- as.data.frame(unlist(residuals_list4)) #make list into dataframe
names(residuals_list4)[1] <- "resid_dev_dur"

#Combine everything into a dataframe
fss_residualtest <- cbind(fss_learning, residuals_list, residuals_list2, residuals_list3, residuals_list4)

#Plot the effect of duration on flow controlling for deviation
partial_plot1 <- ggplot(fss_residualtest, aes(resid_dur_dev, resid_flow_dev)) + geom_point(alpha=.35) + geom_smooth(method="lm", se=FALSE) + 
  facet_wrap("participant") + theme_bw() + ylab("Flow score") + xlab("log(Duration)") +
  labs(title="Controlling for Deviation scores")

#Plot the effect of deviation on flow controlling for deviation
partial_plot2 <- ggplot(fss_residualtest, aes(resid_dev_dur, resid_flow_dur)) + geom_point(alpha=.35) + geom_smooth(method="lm", se=FALSE) + 
  facet_wrap("participant") + theme_bw() + ylab(NULL) + xlab("Deviation scores") +
  labs(title="Controlling for log(Duration)")

ggarrange(partial_plot1, partial_plot2, 
          font.label = list(size = 10, color = "black", face = "bold"), 
          common.legend = TRUE, legend="bottom", hjust = -0.25, ncol = 2, nrow = 1)

ggsave("partial_regression.pdf", width=12, height=7) #uncomment to save figure as pdf

```

## Main analyses controlling for covariates + miscellaneous
```{r Misc, message=FALSE, warning=FALSE}

#Main analyses controlling for covariates
fss_learning_covariates <- cbind(fss_learning, background_flowcar)
fss_learning_covariates2 <- subset(fss_learning_covariates, cumrun!=1)


#f~d effect (only new data; hypothesis 1)
fss_learning_lmer_h1a <- lmer(flow ~ deviation + age + gender + driving_experience + gaming_experience + (deviation|participant), 
                                data=subset(fss_learning_covariates, as.numeric(as.character(ID)) > 9))

#learning effect (only new data; hypothesis 1)
fss_learning_lmer_h1b <- lmer(ln.duration ~ ln.cumrun + age + gender + driving_experience + gaming_experience + (ln.cumrun|participant), 
                                      data=subset(fss_learning_covariates, as.numeric(as.character(ID)) > 9))


#moderating effect of experience (hypothesis 2)
fss_learning_lmer_h2 <- lmer(flow ~ deviation*cumrun + age + gender + driving_experience + gaming_experience + (1|participant), data=fss_learning_covariates)

summary(fss_learning_lmer_h1a)
summary(fss_learning_lmer_h1b)
summary(fss_learning_lmer_h2)

#median, mean, sd, and range of flow (new dataset from 2019)
fss_learning_covariates %>% filter(as.numeric(as.character(ID)) > 9) %>% summarize(mean_flow = mean(flow), 
                                                                     median_flow = median(flow),
                                                                     sd_flow = sd(flow),
                                                                     max_flow = max(flow),
                                                                     min_flow = min(flow))

#median, mean, sd, and range of flow (full dataset)
fss_learning_covariates %>% summarize(mean_flow = mean(flow),
                                      median_flow = median(flow),
                                      sd_flow = sd(flow),
                                      max_flow = max(flow),
                                      min_flow = min(flow))


```
